---
title: "Predicting Elantra Sales"
author: "Terrel Shumway"
date: "04/30/2015"
output: html_document
---

This document presents answers for homework 2 part 5.

## Getting and Cleaning the Data
```{r}
baseurl = "https://courses.edx.org/c4x/MITx/15.071x_2/asset/"

getdata = function(local){
  if(!file.exists(local)){
    library(downloader)
    remote = paste0(baseurl,local)
    print(remote)
    download(remote,local)
  }
  read.csv(local)
}

data = getdata("elantra.csv")

```

## Split for Training and Testing 
problem 1
```{r}
spl = data$Year<2013
train=subset(data,spl)
test=subset(data,!spl)
nrow(train)
```

## Build a model
problem 2.1,2.2:
```{r}
m1 = lm(
  ElantraSales ~ Unemployment + CPI_all + CPI_energy + Queries
  ,data=train)
s = summary(m1)
s$r.squared
```

problem 2.2:
```{r}
sum(s$coefficients[,4]<.05)
```

problem 2.3:
```{r}
s$coefficients["Unemployment",]
```

## Modeling Seasonality

problem 3.1:
```{r}
m2 = lm(
  ElantraSales ~ Month + Unemployment + CPI_all + CPI_energy + Queries
  ,data=train)
s = summary(m2)
s$r.squared
s$coefficients
```

problem 3.3:
```{r}
cm = s$coefficients[["Month","Estimate"]]
cm*(3-1)
cm*(5-1)
```


```{r}
train$mf = as.factor(train$Month)
test$mf = as.factor(test$Month)

m3 = lm(
  ElantraSales ~ mf + Unemployment + CPI_all + CPI_energy + Queries
  ,data=train)
s = summary(m3)
```
problem 4.1: $R^2 = `r s$r.squared`$

problem 4.2:
```{r}
s$coefficients[s$coefficients[,4]<.1,]
```

## Correlations
problem 5.1:
```{r}
cort = cor(train[,c("Month","CPI_all","CPI_energy","Queries")])
e = cort["CPI_energy",]
e[abs(e)>0.6]
```

problem 5.2:
```{r}
q = cort["Queries",]
q[abs(q)>0.6]
```

## A Reduced Model

problem 6.1:
```{r}
m4 = step(m3)
setdiff(attr(m3$terms,"term.labels"),attr(m4$terms,"term.labels"))
```

## Test Predictions


```{r}
pred = predict(m4,newdata=test)
abserr = abs(test$ElantraSales-pred)
sse = sum(abserr^2)
```
problem 6.2: $SSE=`r sse`$

```{r}
baseline = mean(train$ElantraSales)
sst = sum((test$ElantraSales-baseline)^2)
r2 = 1 - sse/sst
wae = which.max(abserr)
memy = paste(test[[wae,"Month"]],test[[wae,"Year"]],sep="/")
```

problem 6.3: $baseline=`r baseline`$

problem 6.4: $R^2 = `r r2`$ 

problem 6.5: `r max(abserr)`

problem 6.6: `r memy`

## Bonus
```{r}
library(ggplot2)
pf = data.frame(my=paste(test$Month,test$Year,sep="/")
                ,predicted=pred,actual=test$ElantraSales)
ggplot(pf,aes(x=predicted,y=actual))+geom_point()+geom_abline(intercept=0, slope=1) +geom_text(label=pf$my,size=4)

```
Interesting: most (`r mean(test$ElantraSales-pred>0)`) of the actual values were higher than predicted (above the line). 



