---
title: "Predicting Parole Violators"
author: "Terrel Shumway"
date: "05/01/2015"
output: html_document
---

This document presents answers for homework 3 part 2.

## Getting and Cleaning the Data
```{r}
library(data.table)

baseurl = "https://courses.edx.org/c4x/MITx/15.071x_2/asset/"

getdata = function(local){
  if(!file.exists(local)){
    library(downloader)
    remote = paste0(baseurl,local)
    print(remote)
    download(remote,local)
  }
  read.csv(local)
}

data = getdata("parole.csv")

data$gender=factor(data$male)
levels(data$gender)=c("female","male")
data$male = NULL

data$race = factor(data$race)
levels(data$race) = c("white","non-white")

data$state = factor(data$state)
levels(data$state) = c("Other","Kentucky","Louisiana","Virginia")

data$crime = factor(data$crime)
levels(data$crime) = c("Other","Larceny","Drug","Driving")
```

## Understanding the Dataset
problem 1.1: `r nrow(data)`
problem 1.2: `r sum(data$violator==1)`

## Splitting the Dataset
```{r}
library(caTools)
set.seed(144)
spl = sample.split(data$violator,SplitRatio=0.7)
train = subset(data,spl)
test = subset(data,!spl)
```

## Building a Model
problem 4.1: 
```{r}
m1 = glm(violator~.,data=train,family="binomial")
s=summary(m1)
s$coefficients[s$coefficients[,4]<.05,]
```



```{r}
logodds = sum(s$coefficients[,1]*c(1 # intercept
                 ,0 # white
                 ,50 # age
                 ,0 # not KY
                 ,0 # not LA
                 ,0 # not VA
                 ,3 # time served
                 ,12 # max time 
                 ,0 # not multiple
                 ,1 # larceny
                 ,0 # not drug
                 ,0 # not driving
                 ,1 # male
                 ))
```

problem 4.2:

$odds = `r exp(logodds)`$

$P(violator) = `r 1/(1+exp(-logodds))`$

## Evaluating the Model on Testing Data

```{r}
library(ROCR)
pred = prediction(predict(m1,type="response"),train$violator)
perf = performance(pred,"tnr","fnr")
plot(perf,colorize=T,print.cutoffs.at=seq(0.2,.6,0.1),text.adj=c(-.3,1.7))

p1 = predict(m1,newdata=test,type="response")
pred = prediction(p1,test$violator)
perf = performance(pred,"tnr","fnr")
plot(perf,colorize=T,print.cutoffs.at=seq(0.2,.6,0.1),text.adj=c(-.3,1.7))
```
problem 5.1: `r max(p1)`


problem 5.2:
```{r}
conf = table(test$violator,p1>0.5)
```

  sensitivity = `r 
  conf[2,2]/sum(conf[2,])
  `
  
  specificity = `r 
  conf[1,1]/sum(conf[1,])
  `
  
  accuracy = `r 
  sum(diag(conf))/length(p1)
  `
  
```{r}
n = nrow(test)
bl.pred = rep(mean(train$violator)>=.5,n)
bl.conf = table(test$violator,bl.pred)
bl.acc = bl.conf[1]/n
```
problem 5.3: `r bl.acc`


problem 5.4,5.5:
```{r}
conf = table(test$violator,p1>0.2)
conf
```

  sensitivity = `r 
  conf[2,2]/sum(conf[2,])
  `
  
  specificity = `r 
  conf[1,1]/sum(conf[1,])
  `
  
  accuracy = `r 
  sum(diag(conf))/length(p1)
  `

problem 5.6:
```{r}
as.numeric(performance(pred,"auc")@y.values)
```

