---
title: "Predicting Loan Repayment"
author: "Terrel Shumway"
date: "05/01/2015"
output: html_document
---

This document presents answers for homework 3 part 3.

## Getting and Cleaning the Data
```{r}
library(data.table)

baseurl = "https://courses.edx.org/c4x/MITx/15.071x_2/asset/"

getdata = function(local){
  if(!file.exists(local)){
    library(downloader)
    remote = paste0(baseurl,local)
    print(remote)
    download(remote,local)
  }
  read.csv(local)
}

data = getdata("loans.csv")

```

## Preparing the Dataset
problem 1.1: `r mean(data$not.fully.paid)`

problem 1.2: 
```{r}
sapply(data,function(x){sum(is.na(x))})
```


```{r}
# storage = "imputed.Rdata.xz"
# if(file.exists(storage)){
#   load(storage)
# }else{
#   library(mice)
#   set.seed(144)
#   measure.vars = setdiff(colnames(data),"not.fully.paid")
#   data[measure.vars]=complete(mice(data[measure.vars]))
#   save(data,file=storage,compress="xz")
# }
data = getdata("loans_imputed.csv")
```

## Splitting the Dataset
```{r}
library(caTools)
set.seed(144)
spl = sample.split(data$not.fully.paid,SplitRatio=0.7)
train = subset(data,spl)
test = subset(data,!spl)
```

## Building a Model
problem 2.1: 
```{r}
m1 = glm(not.fully.paid~.,data=train,family="binomial")
s=summary(m1)
s$coefficients[s$coefficients[,4]<.05,]
```

problem 2.2:
```{r}
logodds = s$coefficients[["fico","Estimate"]]*(700-710)
logodds
exp(logodds)
```

```{r}
test$predicted.risk = predict(m1,newdata=test,type="response")
conf = table(test$not.fully.paid,test$predicted.risk>0.5)
conf
```

```{r}
accuracy = sum(diag(conf))/nrow(test)
sensitivity = conf[2,2]/sum(conf[2,])
specificity = conf[1,1]/sum(conf[1,])
```

```{r}
bl.pred = rep(mean(train$not.fully.paid)>=.5,nrow(test))
bl.conf = table(test$not.fully.paid,bl.pred)
bl.acc = bl.conf[1]/nrow(test)
```

problem 2.3: 

$accuracy = `r accuracy`$

baseline accuracy: $`r bl.acc`$

problem 2.4:

```{r}
library(ROCR)
pred = prediction(test$predicted.risk,test$not.fully.paid)
auc = as.numeric(performance(pred,"auc")@y.values)
```

$AUC = `r auc`$

## A Smart Baseline

problem 3.2:
```{r}
m2 = glm(not.fully.paid ~ int.rate,data=train,family="binomial")
m2.pt = predict(m2,newdata=test,type="response")
max(m2.pt)
```

problem 3.3:
```{r}
m2.auc = auc = as.numeric(performance(prediction(m2.pt,test$not.fully.paid),"auc")@y.values)
m2.auc
```

## Profitability of an Investment
Note: A real profitability analysis would include partial payment and cost of collections. The interest rate would be the dependent variable that the model would set. Therefore, this scenario does not seem very realistic to me. Also, continuous compounding makes the math easy, but I have never seen it used in practice even once.

Problem 4.1
```{r}
10*exp(0.06*3)
```

## A Simple Investment Strategy
```{r}
test$profit = ifelse(test$not.fully.paid,-1,exp(test$int.rate*3)-1)
max(test$profit)*10
```

```{r}
hiint = subset(test,int.rate>=.15)
mean(hiint$profit)
mean(hiint$not.fully.paid)
```

```{r}
cutoff = sort(hiint$predicted.risk,decreasing=FALSE)[100]
basket = subset(hiint,predicted.risk <= cutoff)
profit = sum(basket$profit)
profit
log(profit/3)
table(basket$not.fully.paid)
```


