---
title: "Income Demographics"
author: "Terrel Shumway"
date: "04/27/2015"
output: html_document
---

This document presents answers for part 3 of homework 4.

```{r}
library(caTools)
library(ROCR)
library(rpart)
library(rpart.plot)
library(randomForest)

accuracy = function(conf){sum(diag(conf))/sum(conf)}


baseurl = "https://courses.edx.org/c4x/MITx/15.071x_2/asset/"

getdata = function(local){
  if(!file.exists(local)){
    library(downloader)
    remote = paste0(baseurl,local)
    download(remote,local)
  }
  data = read.csv(local)#,stringsAsFactors=F)
  data
}

data = getdata("census.csv")
```

## A Logistic Regression Model


```{r}
set.seed(2000)
spl = sample.split(data$over50k,SplitRatio=.6)
train=data[spl,]
test=data[!spl,]

```

Problem 1.1
```{r}
storage="over50logreg.Rdata.xz"
if(file.exists(storage)){
  load(storage)
}else{
  m1 = glm(over50k~.,data=train,family="binomial")  
  save(m1,file=storage,compress="xz")
}

summary(m1)

```

Problem 1.2

```{r}
m1.pred = predict(m1,newdata=test,type="response")
m1.conf = table(test$over50k,m1.pred>0.5)
m1.acc = accuracy(m1.conf)
```
model1 accuracy = `r m1.acc`

Problem 1.3

```{r}
base = table(train$over50k)
m0.outcome = dimnames(base)[[1]][which.max(base)]
m0.conf = table(test$over50k,rep(m0.outcome,length=nrow(test)))
m0.acc = accuracy(m0.conf)
```
baseline accuracy = `r m0.acc`

Problem 1.4
```{r}
m1.auc = as.numeric(performance(prediction(m1.pred,test$over50k),"auc")@y.values)
```
area under curve: `r m1.auc`



## A CART Model

```{r}
m2 = rpart(over50k~.,data=train,method="class")
m2.pred = predict(m2,newdata=test)[,2]
m2.conf = table(test$over50k,m2.pred>0.5)
m2.acc = accuracy(m2.conf)
```
Problem 2.1: number of splits = `r m2$cptable[nrow(m2$cptable),"nsplit"]`

Problems 2.2,2.3
```{r}
prp(m2)
```

Problem 2.4: accuracy = `r m2.acc`

Problem 2.5
```{r}

m1.roc = prediction(m1.pred,test$over50k)
plot(performance(m1.roc,"tpr","fpr"),
     colorize=T,print.cutoffs.at=seq(0,1,0.1),text.adj=c(-.3,1.7))

m2.roc = prediction(m2.pred,test$over50k)
plot(performance(m2.roc,"tpr","fpr"),
     colorize=T,print.cutoffs.at=seq(0,1,0.1),text.adj=c(-.3,1.7))

```

Problem 2.6
```{r}
m2.auc = as.numeric(performance(prediction(m2.pred,test$over50k),"auc")@y.values)
```
area under curve: `r m2.auc`

## A Random Forest Model


```{r}
set.seed(1)
trainSmall = train[sample(nrow(train), 2000), ]

set.seed(1)
m3 = randomForest(over50k~.,data=trainSmall)

m3.pred = predict(m3,newdata=test)
m3.conf = table(test$over50k,m3.pred)
m3.acc = accuracy(m3.conf)
```
Problem 3.1: accuracy = `r m3.acc`

Problem 3.2
```{r}
vu = varUsed(m3, count=TRUE)
vusorted = sort(vu, decreasing = FALSE, index.return = TRUE)
dotchart(vusorted$x, names(m3$forest$xlevels[vusorted$ix]))
```

Problem 3.3
```{r}
varImpPlot(m3)
```

## Selecting CP with cross-validation

```{r}

library(caret)
library(e1071)

storage = "cvmodels.Rdata.xz"
if(file.exists(storage)){
  load(storage)
}else{
  set.seed(2)
  numFolds = trainControl(method="cv",number=10)
  cpGrid = expand.grid(.cp = seq(0.002,0.1,0.002))
  cvmodels = train(over50k~., 
        data=train,
        method="rpart",
        trControl=numFolds,
        tuneGrid = cpGrid)
  save(numFolds,cpGrid,cvmodels,file=storage,compress="xz")
}

bestcp = cvmodels$bestTune
```
Problem 4.1: bestcp = `r bestcp`

```{r}
m4 =rpart(over50k~.,data=train,method="class",cp=bestcp)  #cvmodels$finalModel
m4.pred = predict(m4,newdata=test)[,2]
m4.conf = table(test$over50k,m4.pred>0.5)
m4.acc = accuracy(m4.conf)
```
Problem 4.2: accuracy = `r m4.acc`

Bonus:
```{r}
m4.roc = prediction(m4.pred,test$over50k)
plot(performance(m4.roc,"tpr","fpr"),
     colorize=T,print.cutoffs.at=seq(0,1,0.1),text.adj=c(-.3,1.7))

```

Problem 4.3: number of splits = `r m4$cptable[nrow(m4$cptable),"nsplit"]`

