---
title: "Automating Reviews in Medicine"
author: "Terrel Shumway"
date: "04/13/2015"
output: 
  html_document:
    theme: readable
---

This document presents answers to questions in homework 5 part 2.

# Loading the Data

```{r,echo=FALSE}

remote = "https://courses.edx.org/c4x/MITx/15.071x_2/asset/clinical_trial.csv"
local = "clinical_trial.csv"
if(!file.exists(local)){
  library(downloader)
  download(remote,local)
}
data = read.csv(local,stringsAsFactors=F)
```
problem 1.1: `r nrow(data)`

```{r}
data$l1 = nchar(data$abstract)
```

problem 1.2: `r sum(data$l1==0)`

```{r}
data$l2 = nchar(data$title)
```
problemt 1.3: `r data[which.min(data$l2),"title"]`


# Preparing the Corpus

```{r}
library(tm)
library(SnowballC)
pre.corp = function(text){
  corpus = Corpus(VectorSource(text))
  corpus = tm_map(corpus,tolower) 
  corpus = tm_map(corpus,PlainTextDocument)
  corpus = tm_map(corpus,removePunctuation) 
  corpus = tm_map(corpus,removeWords, stopwords("english"))
  corpus = tm_map(corpus,stemDocument)
  corpus
}

corpT = pre.corp(data$title)
dtmT = as.matrix(removeSparseTerms(DocumentTermMatrix(corpT),0.95))

corpA = pre.corp(data$abstract)
dtmA = as.matrix(removeSparseTerms(DocumentTermMatrix(corpA),0.95))
```
problem 2.1: 
title words= `r ncol(dtmT)` abstract words= `r ncol(dtmA)`

## Building a Model

```{r}
mk.df = function(x,prefix){
  y = as.data.frame(x)
  colnames(y) = paste0(prefix,colnames(x))
  rownames(y) = NULL # necessary to suppress spurious warning from cbind
  y
}

dtm = data.frame(trial=data$trial)
dtm = cbind(dtm,mk.df(dtmT,"T"),mk.df(dtmA,"A"))

```

Problem 3.2: `r ncol(dtm)` columns

split into training and testing sets

```{r}
library(caTools)
set.seed(144)
spl = sample.split(dtm$trial,SplitRatio=.7)
train = dtm[spl,]
test = dtm[!spl,]
```

baseline model

```{r}
baseline = ifelse(mean(train$trial)<.5,0,1)
bl.c = table(test$trial)
bl.acc = bl.c[1+baseline]/nrow(test)
```
problem 3.3: accuracy of baseline model=`r bl.acc`

Build a CART model

```{r}
library(rpart)
library(rpart.plot)

m1 = rpart(trial~.,data=train,method="class")
```

problem 3.4:
```{r}
prp(m1)
```

```{r}
m1.pred = predict(m1)
```
problem 3.5: `r max(m1.pred[,2])`


```{r}
N = nrow(train)
m1.pt = predict(m1,type="class")
m1.ct = table(train$trial,m1.pt)
m1.acct = (m1.ct[1,1]+m1.ct[2,2])/N
m1.senst = m1.ct[2,2]/(m1.ct[2,1]+m1.ct[2,2])
m1.spect = m1.ct[1,1]/(m1.ct[1,1]+m1.ct[1,2])
```


problem 3.7:

var                      |value
---------------------    |-----
training set accuracy    |`r m1.acct`
training set sensitivity |`r m1.senst`
training set specificity |`r m1.spect`

## Evaluating the Model on the Testing Set

```{r}
N = nrow(test)
m1.p = predict(m1,newdata=test,type="class")
m1.c = table(test$trial,m1.p)
m1.acc = (m1.c[1,1]+m1.c[2,2])/N
m1.sens = m1.c[2,2]/(m1.c[2,1]+m1.c[2,2])
m1.spec = m1.c[1,1]/(m1.c[1,1]+m1.c[1,2])
```

problem 4.1:

var                      |value
---------------------    |-----
testing set accuracy     |`r m1.acc`
testing set sensitivity  |`r m1.sens`
testing set specificity  |`r m1.spec`


```{r}
library(ROCR)
m1.pred = predict(m1,newdata=test)
m1.roc = prediction(m1.pred[,2],test$trial)
plot(performance(m1.roc,"tpr","fpr")
     ,colorize=T,print.cutoffs.at = seq(0,1,0.1),text.adj=c(-0.2,1.7))
m1.auc = as.numeric(performance(m1.roc,"auc")@y.values)
```
problem 4.2: area under curve (auc) = `r m1.auc`

